---
title: "UpdatedRegression4SNPsAndBodyComp_20191114"
author: "M. Newell"
date: "11/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



Create final dataset and split into male and female dataset
```{r}
#load main dataset
load('/rsgrps/yann/michellenewell/UKB/Full_dataset/Main_dataset_UKB_n487320_derived variables_20191014_mdat.rda')
dim(mdat)
```
Merge SNPs into main file
```{r}
#______________________Merge__________________________________________________________
#Combine 4 SNPs with phenotype data
setwd('/rsgrps/yann/michellenewell/UKB/PheWAS/7Traits_4SNPs_20190909/xData')
snps <- read.csv('TraitsOfInterest_SNPs_7Cardio_n478320_updated5snps_20191108.csv', header = T)

all.equal(snps$userId, mdat$f.eid)
#[1] TRUE

final <- merge(mdat, snps[, c("userId", "rs13389219_T", "snp7_130438531_C", "rs5792311_C", "rs8126001_T", "rs998584_A")], by.x = "f.eid", by.y = "userId")
dim(final)
#[1] 487320   4120
#rs13389219_T and rs998584_A already in data set, merged column from snps dataset ends in '.y'
```
```{r}
library(tidyverse)
```
Create new variables
```{r}
#LegFat %, average of left and right
final2 <- final %>%
  mutate(legFatPerc = ((final$f.23115.0.0 + final$f.23111.0.0)/2))

summary(final2$legFatPerc)
#Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
#1.45   22.25   33.15   31.95   41.00   67.25    8808 

final2$CRP_excl10 <- ifelse((final2$f.30710.0.0 < 10), final2$f.30710.0.0, NA)
summary(final2$CRP_excl10)
```







Create ALT with exclusions for liver disease, certain medications, and high alcohol use

Liver disease
```{r}
#Loop to calculate people with liver illness
#_____________________________Liver disease codes_________________________________________________________________________
X <- c("1155", "1156", "1157", "1158", "1159", "1160", "1161", "1162", "1163", "1164", "1165",
     "1339", "1340", "1047", "1048", "1049", "1050", "1051", "1052", "1053", "1054", "1055", 
     "1056", "1057", "1058", "1578", "1579", "1580", "1581", "1582", "1604", "1439", "1506") 

l1 <- rep(0, nrow(final2))
g1 <- c()

for(i in c(2783:2811)){
  g1 <- ifelse(final2[, i]%in%X, 1, 0)
  l1 <- g1 + l1
  print(i)
}

table(l1)

final2$hepatic_MedConditions <- ifelse(l1 >= 1, 1, 0)
table(final2$hepatic_MedConditions)
```


People on liver medications
```{r}
## amiodarone,prednisolone,methylprednisolone 
Y <-c("1140888502", "1140868364", "1140874976") 


###### loop for liver meds use variable
l2 <- rep(0, nrow(final2))
g2 <- c()

for(i in c(2055:2102)){
  g2 <- ifelse(final2[, i]%in%Y, 1, 0)
  l2 <- g2 + l2
  print(i)
}

table(l2) 

final2$hepatic_exclMeds <- ifelse(l2 >= 1, 1, 0)
table(final2$hepatic_exclMeds)

```
Alcohol Use
```{r}
#_______________________________Alcohol use____________________________________________________________________
#### take average for each across all three instances

final2$drinks_per_wk <- NA

final2$drinks_per_wk <- apply(final2[ , c('f.1568.0.0', 'f.1578.0.0', 'f.1588.0.0', 'f.1598.0.0', 'f.5364.0.0')], 1, sum, na.rm=T)
final2$drinks_per_wk <- ifelse((final2$drinks_per_wk < 0), NA, final2$drinks_per_wk)
```


Create ALT excluding those with liver disease, liver meds, and > 14 alcoholic drinks per day
```{r}
final2$ALT_MultExcl <- NA
final2$ALT_MultExcl <- ifelse((final2$hepatic_MedConditions == 1 | final2$hepatic_exclMeds == 1 | (final2$drinks_per_wk > 14)), NA, final2$f.30620.0.0)
summary(final2$ALT_MultExcl)

```
Stratify by Sex for final datasets
```{r}
female <- final2 %>%
  filter(f.31.0.0 == 0)

male <- final2 %>%
  filter(f.31.0.0 == 1)

```

Then, inverse normalize all body compositions and ALT of interest for female and male
```{r}
#INVERSE NORMALIZATION
my.invnorm = function(x)
{
  res = rank(x)
  res = qnorm(res/(length(res)+0.5))
  return(res)
}

#female, InvNorm BMI
female$InvNorm_BMI[!is.na(female$f.21001.0.0)] <- my.invnorm(female$f.21001.0.0[!is.na(female$f.21001.0.0)])
#female, InvNorm WC
female$InvNorm_WC[!is.na(female$f.48.0.0)] <- my.invnorm(female$f.48.0.0[!is.na(female$f.48.0.0)])
#female, InvNorm HC
female$InvNorm_HC[!is.na(female$f.49.0.0)] <- my.invnorm(female$f.49.0.0[!is.na(female$f.49.0.0)])
#female, InvNorm legfat
female$InvNorm_LegFatPerc[!is.na(female$legFatPerc)] <- my.invnorm(female$legFatPerc[!is.na(female$legFatPerc)])
#female, InvNorm bodyFat
female$InvNorm_BodyFatPerc[!is.na(female$f.23099.0.0)] <- my.invnorm(female$f.23099.0.0[!is.na(female$f.23099.0.0)])
#female, InvNorm ALT
female$InvNorm_ALT_MultExcl[!is.na(female$ALT_MultExcl)] <- my.invnorm(female$ALT_MultExcl[!is.na(female$ALT_MultExcl)])
#male, InvNorm BMI
male$InvNorm_BMI[!is.na(male$f.21001.0.0)] <- my.invnorm(male$f.21001.0.0[!is.na(male$f.21001.0.0)])
#male, InvNorm WC
male$InvNorm_WC[!is.na(male$f.48.0.0)] <- my.invnorm(male$f.48.0.0[!is.na(male$f.48.0.0)])
#male, InvNorm HC
male$InvNorm_HC[!is.na(male$f.49.0.0)] <- my.invnorm(male$f.49.0.0[!is.na(male$f.49.0.0)])
#male, InvNorm legfat
male$InvNorm_LegFatPerc[!is.na(male$legFatPerc)] <- my.invnorm(male$legFatPerc[!is.na(male$legFatPerc)])
#male, InvNorm bodyFat
male$InvNorm_BodyFatPerc[!is.na(male$f.23099.0.0)] <- my.invnorm(male$f.23099.0.0[!is.na(male$f.23099.0.0)])
#male, InvNorm ALT
male$InvNorm_ALT_MultExcl[!is.na(male$ALT_MultExcl)] <- my.invnorm(male$ALT_MultExcl[!is.na(male$ALT_MultExcl)])
#female, InvNorm CRP
female$InvNorm_CRP[!is.na(female$CRP_excl10)] <- my.invnorm(female$CRP_excl10[!is.na(female$CRP_excl10)])
#male, InvNorm CRP
male$InvNorm_CRP[!is.na(male$CRP_excl10)] <- my.invnorm(male$CRP_excl10[!is.na(male$CRP_excl10)])
```

Diagnostic plots
```{r}
#Female, BMI, VEGFA
reg1 <- lm(InvNorm_BMI ~ rs998584_A.y + Age_v0 + Age_sq + as.factor(female$f.54.0.0) + as.factor(female$chip) + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10, data = female)
summary(reg1)
plot(reg1)
```
```{r}
#Female, GRB14, WC
reg2 <- lm(InvNorm_WC ~ rs13389219_T.y + Age_v0 + Age_sq + as.factor(female$f.54.0.0) + as.factor(female$chip) + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10, data = female)
summary(reg2)
plot(reg2)
```
```{r}
#Male, GRB14, HC
reg3 <- lm(InvNorm_HC ~ rs13389219_T.y + Age_v0 + Age_sq + as.factor(male$f.54.0.0) + as.factor(male$chip) + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10, data = male)
summary(reg3)
plot(reg3)
```
```{r}
#Male, RGS19, Body Fat
reg4 <-lm(InvNorm_BodyFatPerc ~ rs8126001_T + Age_v0 + Age_sq + as.factor(male$f.54.0.0) + as.factor(male$chip) + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10, data = male)
summary(reg4)
plot(reg4)
```
```{r}
#male, KLF14, ALT
reg5 <- lm(InvNorm_ALT_MultExcl ~ snp7_130438531_C + Age_v0 + Age_sq + as.factor(male$f.54.0.0) + as.factor(male$chip) + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10, data = male)
summary(reg5)
plot(reg5)
```
```{r}
#Female, KLF14, legfat
reg6 <- lm(InvNorm_LegFatPerc ~ snp7_130438531_C + Age_v0 + Age_sq + as.factor(female$f.54.0.0) + as.factor(female$chip) + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10, data = female)
summary(reg6)
plot(reg6)
```

```{r}
#Female, CRP, VEGFA
reg7 <- lm(InvNorm_CRP ~ rs998584_A.y + Age_v0 + Age_sq + as.factor(female$f.54.0.0) + as.factor(female$chip) + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10, data = female)
summary(reg7)
plot(reg7)
```
```{r}
#Male, CRP, Body Fat
reg8 <-lm(InvNorm_CRP ~ rs8126001_T + Age_v0 + Age_sq + as.factor(male$f.54.0.0) + as.factor(male$chip) + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10, data = male)
summary(reg8)
plot(reg8)
```


j = c("InvNorm_BMI", "InvNorm_WC", "InvNorm_HC", "InvNorm_LegFatPerc", "InvNorm_BodyFatPerc", "InvNorm_ALT_MultExcl")
k = c("rs13389219_T.y", "snp7_130438531_C", "rs8126001_T", "rs998584_A.y")
covariates_f <- "Age_v0 + Age_sq + as.factor(female$f.54.0.0) + as.factor(female$chip) + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10"
covariates_m <- "Age_v0 + Age_sq + as.factor(male$f.54.0.0) + as.factor(male$chip) + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10"






